---
title: "Cross-linguistic variation in early word learning"
author: "Daniel Yurovsky, Mika Braginsky, Virginia Marchman, & Michael C. Frank"
date: "2015-05-08"
output:
  html_document:
  highlight: tango
theme: spacelab
---

Baby's First 10 Words Cross-linguistic

```{r setup, include=FALSE}
library(knitr)
options(digits = 2)
opts_chunk$set(message=FALSE, warning=FALSE, cache=TRUE)
```

Load required libraries.
```{r librares, cache=FALSE}
library(dplyr)
library(directlabels)
library(RMySQL)
library(tidyr)
library(bootstrap)
library(ggplot2)
library(RCurl)
library(magrittr)
```

Get a script that provides interface functions for pulling data out of Wordbank.
```{r wordbank.funs, cache=FALSE}
url <- 'https://raw.githubusercontent.com/langcog/wordbank/master/shiny_apps/data_loading.R'
script <- getURL(url, ssl.verifypeer = FALSE)
eval(parse(text = script))
```

Connect to the Wordbank database.
```{r connect, cache=FALSE}
wordbank <- connect.to.wordbank("prod")
```

Load tables
```{r tables}
common.tables <- get.common.tables(wordbank)
instrument.tables <- get.instrument.tables(wordbank, common.tables)

admins <- get.administration.data(common.tables)
items <- get.item.data(common.tables)
```

Filter down to appropriate kids
```{r}
vocab.admins <- admins %>%
  select(data_id, language, form, age, sex, production) %>%
  filter(form == "WG", !is.na(sex), sex != "", age >= 8, age <=16)
```

Function that get's one language's data from wordbank
```{r}
get.language.data <- function(lang, vocab.data) {
  
  lang.table <- filter(instrument.tables, language==lang, form=="WG")$table[[1]]
  
  words <- items %>%
    filter(type == "word", language == lang, form == "WG") %>%
    select(definition,item.id,uni_lemma,category) %>%
    rename(item_id = item.id)
  
  lang.data<- lang.table %>%
    filter(basetable_ptr_id %in% vocab.data$data_id ) %>%
    select_(.dots=c("basetable_ptr_id",words$item_id)) %>%
    as.data.frame %>% 
    gather(item_id,value,-basetable_ptr_id) %>%
    rename(data_id = basetable_ptr_id) %>%
    left_join(vocab.data)
  
  return(select(left_join(lang.data,words),-item_id))

  }
```


Look at age distributions
```{r,fig.width=3,fig.height=8}
# all.by.age <- vocab.admins %>%
#   group_by(language,age) %>%
#   summarise(all.n = n())
# 
# age.hists <- vocab.data %>%
#   group_by(language,age) %>%
#   summarise(n = n()) %>%
#   left_join(all.by.age) %>%
#   mutate(prop = n/all.n) %>%
#   mutate(prop = prop/sum(prop))
# 
# ggplot(data = age.hists,aes(x=age,y=prop)) +
#   facet_grid(language ~ .) +
#   geom_histogram(stat="identity",
#                  fill="steelblue") +
#   geom_vline(xintercept=12,linetype="dashed")+
#   scale_x_continuous(limits=c(8,16),breaks=seq(8,16),name="Age (months)") +
#   scale_y_continuous(name = "Proportion of Children") +
#   theme_bw()+
#   theme(panel.grid=element_blank()) 
```

Get all data
```{r}
languages <- c("English","Norwegian")
all.data <- bind_rows(sapply(languages, 
                             function(lang) get.language.data(lang,vocab.admins),
                             simplify = FALSE))


production.data <- all.data %>%
  ungroup() %>%
  mutate(language = factor(language)) %>%
  group_by(language,category,uni_lemma,definition) %>%
  summarise(prop = mean(value == "produces",na.rm=TRUE)) %>%
  summarise(prop = max(prop,na.rm=TRUE)) %>%
  spread(language,prop) %>%
  filter(!is.na(English),!is.na(Norwegian)) %>%
  gather(language,prop,English:Norwegian) %>%
  group_by(language) %>%
  arrange(desc(prop)) %>%
  mutate(order = 1:n()) %>%
  rowwise() %>%
  mutate(comb.lemma = paste0(category,"-",uni_lemma))

cors <- rep(0,length(unique(production.data$comb.lemma)))

for(i in 1:length(cors)) {
  eng.orders <-  filter(production.data,language=="English",order <= i)
  nor.orders <- filter(production.data,language=="Norwegian",
                       comb.lemma %in% unique(eng.orders$comb.lemma))
  
  diff <- bind_rows(eng.orders,nor.orders) %>%
    group_by(comb.lemma) %>%
    summarise(diff = max(order) - min(order)) %>%
    summarise(diff = mean(diff))
  
  cors[i] <- diff[[1]]
}

cors <- data.frame(order=1:length(cors),distance=cors)
```


```{r}
quartz(width=6,height=4)
ggplot(data=cors,aes(x=order,y=distance)) +
  geom_point()+
  theme_bw()+
  theme(panel.grid=element_blank())
```




kids.by.lang <- vocab.data %>%
  group_by(language) %>%
  summarise(num.kids = n())

produced.data <- all.data %>%
  filter(value == "produces") %>%
  group_by(language,category,definition,gloss) %>%
  summarise(mean.age = mean(age),
            n = n()) %>%
  group_by(language) %>%
  left_join(kids.by.lang) %>%
  mutate(prop = n/num.kids) %>%
  arrange(desc(prop))
```

Top words
```{r}
top.words <- produced.data %>%
  slice(1:10) %>%
  select(language,gloss,prop,definition)

tab.words <- top.words %>%
  mutate(order = 1:10) %>%
  rowwise() %>%
  mutate(word = paste0(gloss," (",round(prop,digits=2),")")) %>%
  select(language,word,order) %>%
  spread(language,word) %>%
  select(-order)

kable(tab.words)
```

Models
```{r,}
model.words <- items %>%
  select(definition,language,gloss,category) %>%
  filter(language %in% unique(produced.data$language)) %>%
  left_join(produced.data) %>%
  rowwise() %>%
  mutate(n = as.numeric(ifelse(is.na(n),0,n)),
         prop = as.numeric(ifelse(is.na(prop),0,prop)),
         len = nchar(definition))

outputs <- NULL
components = c("count","zero")
params = c("len")

languages <- unique(model.words$language) 

predict.params <- expand.grid(language = languages,
                              component = components,
                              param = params)%>%
  arrange(language,component,param)

for(lang in languages) {
  
  hurd <- hurdle(n ~ len, 
                 data = filter(model.words,language==lang))
  


  for(component in components) {

    model.outs <- summary(hurd)$coefficients[as.character(component)][[1]]
    
    outputs <- rbind(outputs,model.outs[c("len"),
                                        c("Estimate", "Std. Error","Pr(>|z|)")])
  }
    
}
colnames(outputs) = c("estimate","se","p")

predict.params <- cbind(predict.params,outputs) %>%
  mutate(ci = 1.96*se) %>%
  mutate(component = factor(component, labels = c("Count Estimate", "Hurdle")))
```

Plot parameters
```{r, fig.width=5.5, fig.height=5}
ggplot(data = predict.params, 
       aes(x = language, y = estimate, fill=component))+
  geom_histogram(stat="identity",position="identity")+
  geom_linerange(aes(ymax =estimate+ci,
                      ymin = estimate-ci)) +
  facet_grid(component ~ param)+
  ylab("Parameter Estimate (+/- 95% CI)")+
  xlab("Dataset") + 
  geom_hline(yintercept=0, lty=2,size=.7) + 
  #scale_color_brewer(name="Dataset",palette="Set1")+
  scale_fill_brewer(palette="Set1")+
  theme_bw(base_size=14) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=-45, hjust = 0),
        axis.title.x = element_text(vjust=-0.5),
        panel.grid = element_blank())
```


